---
title: "Stratification of the Late-Middle-Age Moravian nobility according to dowry, replication material"
author: "Michael Skvrnak"
date: "June 22, 2016"
output: html_document
---

## Loading data
```{r}
library(magrittr)

vena_basic <- read.csv("vena_basic.csv", 
                       stringsAsFactors = FALSE)

vena_basic$sum <- as.numeric(vena_basic$sum)

head(vena_basic)
```

###Description of the data:

Variables:

* __ID__ - identification number of the dower
* __rok__ - year of the record
* __idd__ - identification number of the dynasty
* __pos__ - status of the recipient (NA - unknown, 0 - high nobility, 
1 - pauperized high nobility, 2 - lower nobility)
* __idd.1__
* __pau__ - status of the donor (0 - high nobility, 
1 - pauperized high nobility)
* __sum__ - the value of the dower
* __vec__ - widower (1)/non-widower(0)
* __va__ - widow (1)/non-widow(0)
* __addtrans__ - no changes to the dower (0) / adding to the dower (1) / transfer of the dower to different assets (2)

## Ratio between dower and dowry
```{r}
library(ggplot2)

# complete observations only
dowry <- c(200, 1000, 250, 200, 350, 600, 100, 200, 200)
dower <- c(260, 1000, 250, 500, 500, 1000, 150, 200, 500)
coefs <- dower/dowry

#mean(coefs)
summary(coefs)

df <- data.frame(dowry = dowry, dower = dower, coefs = coefs)

ggplot(data = df, aes(x = coefs)) + 
  geom_histogram(bins = 10, 
                 color = "black", 
                 fill = "white") + 
  xlab("Ratio between dower and dowry") + 
  ylab("Number of observations") + 
  ggtitle("Figure 1: Ratio between dower and dowry")


```


*"According to Vsehrd, the dower should be 2.5 times higher than the dowry"*

That does not hold true. Although we know both the values of dower 
and dowry only in nine cases, the dower reached the ratio of 2.5 only 
in 2 of them. In the rest of the observations, it did not reach
even 2 and the most common ratio was 1. 


## EDA: Dowers in time 

```{r}

dar <- ggplot(data = vena_basic, aes(x = rok, y = sum, color = factor(pau))) + geom_jitter() + 
  labs(x = "Year", y = "Value of the dower", color = "Status of donor") + theme(legend.position = "bottom") + 
  scale_color_manual(values = c("black", "red"), 
                     labels = c("High nobility", 
                              "Pauperized high nobility")) + 
  ggtitle("Figure 3: Dower by the status of donors")
dar

prijem <- ggplot(data = vena_basic, aes(x = rok, y = sum, color = factor(pos)))
prijem + geom_jitter() + 
  labs(x = "Year", y = "Value of the dower", 
       color = "Status of recipient") + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = c("grey", "blue", "green", "red"), 
                     labels = c( "Unknown", "High nobility", 
                               "Pauperized nobility",
                               "Lower nobility")) + 
  ggtitle("Figure 4: Dower by the status of recipient")


vena_basic$addtrans[vena_basic$addtrans == "?"] <- NA

trans <- ggplot(data = vena_basic, 
                aes(x = rok, y = sum, 
                    color = factor(addtrans)))
trans + geom_jitter() + 
  labs(x = "Year", y = "Value of the dower", 
       color = "Changes in dower") + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = c("blue", "green", "red"), 
                     labels = c("Regular dower", 
                              "Adding to the dower", 
                              "Transfer of the dower")) + 
  ggtitle("Figure 5: Adding to and transferring of dowers")

vdovec <- ggplot(data = vena_basic, 
                 aes(x = rok, y = sum, color = factor(vec)))
vdovec + geom_jitter() + 
  labs(x = "Year", y = "Value of the dower", color = "") + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("Non-widower", "Widower")) + 
  ggtitle("Figure 6: Dowers given by  widowers")

vdovy <- ggplot(data = vena_basic, 
                aes(x = rok, y = sum, color = factor(va)))
vdovy + geom_jitter() + 
  labs(x = "Year", y = "Value of the dower", color = "") + 
  theme(legend.position = "bottom") + 
  scale_color_manual(values = c("blue", "red"), 
                     labels = c("Non-widow", "Widow")) + 
  ggtitle("Figure 7: Dowers received by widows")
```


## Charts 2: Size of the dower
```{r}
#plot <- ggplot(data=vena_basic, aes(x=sum, y=pau))
#plot + geom_point()

# Size of the dower by the status of the donor
plot2 <- ggplot(data = vena_basic, aes(x = factor(pau), y = sum))
plot2 + geom_jitter() + 
  geom_boxplot(color = "red", fill = NA, outlier.colour = NA) + 
  coord_flip() + 
  labs(y = "Value of the dower", x = "Status of the donor") 

# Size of the dower by the status of the recipient
plot3 <- ggplot(data = vena_basic, aes(x = factor(pos), y = sum))
plot3 + geom_jitter() + 
  geom_boxplot(color = "red", fill = NA, outlier.colour = NA) + 
  coord_flip() + 
  labs(y = "Value of the dower", x = "Status of the recipient")

hist2 <- ggplot(data = vena_basic, aes(x = sum))
hist2 + geom_histogram(color = "black", fill = "white") + 
  labs(x = "Value of the dower", y = "Number of cases") + 
  ggtitle("Figure 2: Size of dowers")

summary(vena_basic$sum)
```



## Correlations
```{r}
# Correlation between dower and status of the donors
cor(vena_basic$sum[!is.na(vena_basic$sum == TRUE)], 
    vena_basic$pau[!is.na(vena_basic$sum == TRUE)], 
    method = "pearson")

# Correlation between dower and status of the recipient 
# (excluding the unknown dynasties)
kor2 <- vena_basic[vena_basic$pos != "?", ]
kor2$pos <- as.numeric(kor2$pos)
cor(kor2$pos[!is.na(kor2$sum == TRUE)], 
    kor2$sum[!is.na(kor2$sum == TRUE)], 
    method = "spearman")

```

The correlation coefficient indicate that both the status of the donor
and the recipient influenced the size of the dower. In particular, 
higher status led to higher dower.



```{r}
# Who married who (absolute numbers)
table(vena_basic$pau, vena_basic$pos)


# Who married who (row percent)
prop.table(table(vena_basic$pau, vena_basic$pos), 1)

```
The tables show that the groom from high nobility married
predominantly brides from high nobility. However, grooms from
pauperized dynasties had higher chances to marry a bride from
higher nobility than the other way around. 

## The regression model of dower 

Hypotheses:

H1: Higher the status of the donor, the higher the value of the
dower.

H2: Higher the status of the recipient, the higher the value of
the dower.

H3: Widows received lower dowers than non-widows.

H4: Widowers gave lower dower than non-widowers because it is 
expected that they have children from previous marriage.

H5: Dowers to which were added later were lower than the rest
(because they were given by young noblemen who had not much
capital)

```{r}
# Recode the status of the recipient into two categories
## 0 = high nobility
## 1 = the rest (pauperized, lower, unknown dynasties)
vena_basic$pos[vena_basic$pos != 0] <- 1

model1 <- lm(log1p(sum) ~ pau + factor(pos), data = vena_basic)
summary(model1)
plot(model1)

model2 <- update(model1, . ~ . + va)
summary(model2)

model3 <- update(model2, . ~ . + vec)
summary(model3)

## Recode the addtrans variables so that:
## 1 = adding to the dower
## 0 = the rest (no change, transfer of the dower)
vena_basic$addtrans[vena_basic$addtrans != 1] <- 0

model4 <- update(model3, . ~ . + factor(addtrans))
summary(model4)

```

